---
title: 'H9 HA216 in vivo competition quail infection: whole genome mutation distribution'
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Analysis of alternative variants distribution across Influenza A genome. This analysis evaluates the  position in the genome divided by Influenza A segments, nature (synonymous, nonsynonymous, nonsense and variants present in the UTRs) along with the frequency of the variants.

The inputs required for this analysis are:

1. Quail.Complete.txt
  This table has sample, quail number, exposure group (inoculated or contact) and group     (var, varL, varLQ, varQ)

2. Complete.Extended.AF.Cov.final.tsv
  This table is a concatenated version that has the information from the variant analysis using LoFreq under GATK guidelines. Variants in the file have at least a frequency of 0.02 and a depth of coverage  of 400X.
  
 3. Complete.SNPdat.Ext.txt
 This table is the result of concatenating outputs produced by [SNPdat](https://code.google.com/archive/p/snpdat/) for each sample evaluated in the study. This table has the type of mutation that each variant represents.
 
 NOTE: Extended and Ext in the names of the files refers to the group of quails that have been use to perform this analysis. This group correspond only to quails that were evaluated throughout the complete experiment. This distinction is important since some animals (=3/group/Exposure and not included in the "extended" group) were euthanize at 5 dpi/dpc for tissue collection.
```{r message=F}
library(dplyr) 
library(ggplot2)
library(stringr)
```

```{r}
## Import files and add headers
## Import Quail.Complete.txt
Quail.Complete <- read.table("Quail.Complete.txt", 
                             header = F, sep = "\t", stringsAsFactors = T)
colnames(Quail.Complete) <- c("Sample", "Quail", "day", "Expo", "Group")
Quail.Complete$Sample <- as.character(Quail.Complete$Sample)
##Erase trailing space in column "Samples"
Quail.Complete$Sample <-trimws(Quail.Complete$Sample,"r")

## Import Complete.Extended.AF.Cov.final.tsv
Complete.Extended.AF.Cov.final <-read.table("Complete.Extended.AF.Cov.final.tsv", head = F)
colnames(Complete.Extended.AF.Cov.final) <- 
  c("Segment", "Position", "Cov", "Length", "REF", "ALT", "Freq", "Sample")

## Complete.SNPdat.Ext.txt
Complete.SNPdat.Ext <- read.table("Complete.SNPdat.Ext.txt", 
                                  sep = "\t", stringsAsFactors=F, head = F, na.strings=c("","NA"))
colnames(Complete.SNPdat.Ext) <- 
  c("Segment", "Position", "REF-AA", "ALT-AA", "SYN", "Sample")

# Erase trailing spaces
Complete.SNPdat.Ext$SYN <-trimws(Complete.SNPdat.Ext$SYN)
# Replace "" for "O"
Complete.SNPdat.Ext$SYN[Complete.SNPdat.Ext$SYN ==""] <- "O"

```

Join tables into one.

```{r}
## Join(s)

# Complete.SNPdat.Ext have duplicated entries based on synonymous mutations found 
# in the same positions within the same segment
Complete.SNPdat.Ext <- Complete.SNPdat.Ext %>% 
  unique()

# Convert columns to the same class
Complete.SNPdat.Ext$Position <- 
  as.character(Complete.SNPdat.Ext$Position)

Complete.Extended.AF.Cov.final$Segment <- 
  as.numeric(Complete.Extended.AF.Cov.final$Segment)

Complete.Extended.AF.Cov.final$Position <- 
  as.character(Complete.Extended.AF.Cov.final$Position)

Complete.SNPdat.Ext$Segment <- 
  as.numeric(Complete.SNPdat.Ext$Segment)

Complete.Extended.AF.Cov.final$Sample <-
  as.character(Complete.Extended.AF.Cov.final$Sample)

# Join
Complete.Extended.AF.Cov.SNPdat <- Complete.Extended.AF.Cov.final %>%
  left_join(Complete.SNPdat.Ext, by = c("Sample", "Segment", "Position")) 

# Add Quail information
Complete.Extended.AF.Cov.SNPdat.final <- Complete.Extended.AF.Cov.SNPdat %>%
  left_join(Quail.Complete, by = c("Sample")) 

# Erase trailing blank characters
Complete.Extended.AF.Cov.SNPdat.final$Expo <- 
  trimws(Complete.Extended.AF.Cov.SNPdat.final$Expo)
Complete.Extended.AF.Cov.SNPdat.final$Group <- 
  trimws(Complete.Extended.AF.Cov.SNPdat.final$Group)

```

In previous analysis, time of exposure for both inoculated and contact groups were as "dpi". The following chunk converts dpi to dpc for contact groups. This chunk also removes 9 dpi samples since there are not enough animals to compare to previous time points.

```{r}
#########                                           #############
# Fix annotation discrepancies and time points with fewer genomes#
########                                            #############

## Replace DC day dpi annotation for dpc annotation
# Create list
q = c("3" = "1",
      "5" = "3",
      "7" = "5",
      "9" = "7",
      "10" = "8")

# Substitute        
Complete.Extended.AF.Cov.SNPdat.final<-Complete.Extended.AF.Cov.SNPdat.final %>% 
  mutate(day = case_when(Expo == "DC" ~ str_replace_all(day, q),
                         TRUE ~ as.character(day)))

## Remove 9 dpi
Complete.Extended.AF.Cov.SNPdat.final <-
  Complete.Extended.AF.Cov.SNPdat.final %>% 
  filter(!(Expo == "DI" & day == 9)) 

## Flag nonsense mutations
Complete.Extended.AF.Cov.SNPdat.final <-
Complete.Extended.AF.Cov.SNPdat.final %>% 
  mutate(SYN = 
           case_when((is.na(`ALT-AA`) & SYN == "N") ~ str_replace(SYN, "N", "X"), 
                         TRUE ~ as.character(SYN)))

write.table(Complete.Extended.AF.Cov.SNPdat.final, 
            file = "Complete.Extended.AF.Cov.SNPdat.final.txt", 
            quote = F, sep="\t",
            row.names = FALSE)
```

As plotting both exposure groups together makes the plot busy, these groups were split.

```{r}

## Filter DI and DC into SYN (Y), NONSYN (N), No.cod (O) and STOP (X)

## ----- DI
DI.Extended.AF.Cov.SNPdat.final.SYN <- 
  Complete.Extended.AF.Cov.SNPdat.final %>%
  filter(SYN == "Y" &  Expo == "DI") 

DI.Extended.AF.Cov.SNPdat.final.NONSYN <- 
  Complete.Extended.AF.Cov.SNPdat.final %>%
  filter(SYN == "N" &  Expo == "DI") 

DI.Extended.AF.Cov.SNPdat.final.No.cod <- 
  Complete.Extended.AF.Cov.SNPdat.final %>%
  filter(SYN == "O" &  Expo == "DI") 

DI.Extended.AF.Cov.SNPdat.final.STOP <- 
  Complete.Extended.AF.Cov.SNPdat.final %>%
  filter(SYN == "X" &  Expo == "DI") 

## ----- DC
DC.Extended.AF.Cov.SNPdat.final.SYN <- 
  Complete.Extended.AF.Cov.SNPdat.final %>%
  filter(SYN == "Y" &  Expo == "DC") 

DC.Extended.AF.Cov.SNPdat.final.NONSYN <- 
  Complete.Extended.AF.Cov.SNPdat.final %>%
  filter(SYN == "N" &  Expo == "DC") 

DC.Extended.AF.Cov.SNPdat.final.No.cod <- 
  Complete.Extended.AF.Cov.SNPdat.final %>%
  filter(SYN == "O" &  Expo == "DC") 

DC.Extended.AF.Cov.SNPdat.final.STOP <- 
  Complete.Extended.AF.Cov.SNPdat.final %>%
  filter(SYN == "X" &  Expo == "DC") 

```

The following will generate the plot SuppFigure 2.

```{r}
## DC and DI plots are made separately
### Plot
# Assign colors
groupColors = c("var" = "red",
                "varL" = "gold",
                "varLQ" = "green3",
                "varQ" = "dodgerblue2")

Expo.label <- c("DI" = "Inoculated",
                "DC" = "Contacts")

DC.Extended.AF.Cov.SNPdat.plot <- 
    ggplot() + 
  geom_point(data = DC.Extended.AF.Cov.SNPdat.final.SYN, 
             aes(x=Length, y=Freq, fill = Group),
             shape=21, size = 2) + 
  geom_point(data = DC.Extended.AF.Cov.SNPdat.final.NONSYN, 
             aes(x=Length, y=Freq, fill = Group), 
             shape =24, size = 2) +
  geom_point(data = DC.Extended.AF.Cov.SNPdat.final.No.cod, 
             aes(x=Length, y=Freq, fill = Group), 
             shape =23, size = 2) +
  geom_point(data = DC.Extended.AF.Cov.SNPdat.final.STOP, 
             aes(x=Length, y=Freq, fill = Group), 
             shape =22, size = 2) +
  scale_fill_manual(values = groupColors) +
  geom_hline(yintercept = 0.5, colour= "grey50", linetype="dashed") +
  geom_hline(yintercept = 1, linetype="dashed") +
  scale_x_continuous(breaks = c(1, 2341, 4682, 6915, 8658, 
                                10223, 11683, 12710, 13600)) +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_blank(), 
    axis.line = element_line(colour = "black"),
    axis.text.x = element_blank(), 
    axis.text.y = element_text(face="bold", size=25)
  ) +
  scale_y_log10() +
  facet_grid(day ~ Expo, labeller = labeller(day = label_value, 
                                 Expo = Expo.label)) + 
  theme_bw() + 
  ylab("Frequency") + xlab("Segments")
print(DC.Extended.AF.Cov.SNPdat.plot)
ggsave("DC.Extended.AF.Cov.SNPdat.pdf", width=10, height=12.5, dpi=300)
```

This figure was further modified using [Inkscape](https://inkscape.org) for publication.
