---
title: 'H9 HA226 in vivo competition quail infection: Common Variants'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

The input required for this analysis is "Complete.Extended.AF.Cov.SNPdat.final.txt" from 'H9 HA226 in vivo competition quail infection: whole genome mutation distribution' 

Here we evaluate variant detection by quail within exposure group


```{r message=F}
library(dplyr)
library(ggplot2)

Complete.Extended.AF.Cov.SNPdat.final <-
  read.table("Complete.Extended.AF.Cov.SNPdat.final.txt", 
                                                   head = T)

# Filter occurrences that are present > 1
# Convert Quails' values to character
Complete.Extended.AF.Cov.SNPdat.final$Quail <-
as.character(Complete.Extended.AF.Cov.SNPdat.final$Quail)

Complete.VariantShared.Ext <-
  Complete.Extended.AF.Cov.SNPdat.final %>%
  group_by(Expo,
           Group, 
           Quail,
           Segment,
           Position,
           Length) %>%
  filter(n()>1) %>%
  count() 
```

Variant detection by quail within exposure group

```{r}

# Count occurrences in Length for df Complete.VariantShared.Ext. 
Quails.VariantShared.Ext <-
Complete.VariantShared.Ext %>%
  group_by(Expo, 
           Segment, 
           Position, 
           Length) %>%
  filter(n()>1) %>%
  count()

Quail.colors <- colorRampPalette(c("light blue", "red"))  (9)

Quails.VariantShared.Ext.plot <-
ggplot(Quails.VariantShared.Ext, aes(x=Length, y= Expo, fill=factor(n))) + 
  geom_tile(width=20) +
  #geom_text(aes(label=Position), angle = 45) + 
  #The above line allowed to number positions in the figure
  scale_fill_manual(values = Quail.colors) +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_blank(), 
    axis.line = element_line(colour = "black"),
    axis.text.y = element_text(size=12),
    axis.text.x = element_blank(),
    strip.text.y = element_text(size = 9), 
    axis.title = element_text(size = 12),
    legend.title.align=0.5
  ) +
  scale_x_continuous(breaks = c(1, 2341, 4682, 6915, 
                                8658, 10223, 11683, 
                                12710, 13600), 
                     limits = c(1, 13600)) + 
  # The above scale divides the segments.
  xlab("Segments") + ylab("Exposure group") +
  labs(fill = "Number\nof\nQuails")
print(Quails.VariantShared.Ext.plot)  
ggsave("Quails.VariantShared.Ext.plot.pdf", width=10, height=3, dpi=300)
```

Image in the publication was further modified for aesthetics purposes using [Inkscape](https://inkscape.org).



